---
title: "Point Process Simulation"
output: pdf_document
---

```{r echo=FALSE}
want = c("MASS", "mvtnorm", "tidyverse", "xtable", "MCMCpack", "doParallel", "compositions", "coda")
have = want %in% rownames(installed.packages())
if ( any(!have) ) { install.packages(want[!have], quiet = TRUE, verbose = FALSE) }

for (lib in want) {
  suppressMessages(library(lib, character.only = TRUE, quietly = TRUE))
}
registerDoParallel(cores = detectCores() - 1)
options(xtable.comment = FALSE)
knitr::opts_chunk$set(cache = T)

```


# Estimated Parameters
```{r}
set.seed(1899)
ad_types <- c("search", "display", "other", "purchase")
K <- length(ad_types)
alpha_mean <- matrix(c(2.817, 0.0860, 0.5381, 0.6167,
                      0.1614, 1.7818, 0.2055, 0.0845,
                      0.4647, 0.1270, 8.0526, 0.8384),
                     nrow = 3, ncol = 4, byrow = T)
row.names(alpha_mean) <- ad_types[1:3]
colnames(alpha_mean) <- ad_types

alpha_sd <- matrix(c(0.1765, 0.0214, 0.0562, 0.0633,
                      0.0496, 0.2314, 0.0572, 0.0347,
                      0.0654, 0.0367, 0.4117, 0.0867), 
                   nrow = 3, ncol = 4, byrow = T)
row.names(alpha_sd) <- ad_types[1:3]
colnames(alpha_sd) <- ad_types

beta_mean <- c(34.0188, 46.8854, 51.5114)
beta_sd <- c(1.7426, 4.9370, 2.3241)
names(beta_mean) <- ad_types[1:3]

theta_mu_mean <- c(-5.3926, -6.1027, -5.8063, -9.7704) 
theta_mu_sd <- c(0.0166, 0.0212, 0.0221, 0.0762) 
names(theta_mu_mean) <- ad_types
names(theta_mu_sd) <- ad_types

Sigma_mu_mean <- matrix(c(0.4584, -0.1197, -0.4942, 0.2256,
                     -0.1197, 0.5934, -0.3380, -0.4157,
                     -0.4942, -0.3380, 1.0014, 0.0762,
                     0.2256, -0.4157, 0.0762, 2.3914), nrow = 4, ncol = 4)

Sigma_mu_sd <- matrix(c(0.0246, -0.0212, 0.0257, 0.1228,
                     0.0212, 0.0335, 0.0304, 0.1665,
                     0.0257, 0.0304, 0.0365, 0.2440,
                     0.1228, 0.1665, 0.2440, 0.2575), nrow = 4, ncol = 4)

row.names(Sigma_mu_mean) <- ad_types
colnames(Sigma_mu_mean) <- ad_types
  
n <- 12000
#inv_df <- n - 1
inv_df <- n - 1
psi_mean <- c(-0.5664, -0.7556, -0.6235, 0.2787)
psi_sd <- c(0.1228, 0.2348, 0.1229, 0.2160)
names(psi_mean) <- ad_types
names(psi_sd) <- ad_types


first_click_prob <- c(1199, 418, 811) %>% prop.table()
```

```{r eval=FALSE, results = 'asis', echo=F}
knitr::kable(alpha_mean, caption = "alpha", escape = F)
```

```{r eval=FALSE, results = 'asis', echo=F}
knitr::kable(beta_mean, caption = "beta", escape = F)
```

```{r eval=FALSE, results = 'asis', echo=F}
knitr::kable(theta_mu_mean, caption = "theta", escape = F)
```

```{r eval=FALSE, results = 'asis', echo=F}
knitr::kable(Sigma_mu_mean, caption = "Sigma", escape = F)
```

```{r eval=FALSE, results = 'asis', echo=F}
knitr::kable(psi_mean, caption = "psi", escape = F)
```

# Simulate Data

## Draw Parameters
1. Draw $\alpha, \beta, \psi, \theta_{\mu}, \Sigma_{\mu}$ and generate $\mu^i \sim MVN_K(\theta_{\mu}, \Sigma_{\mu})$ to simulate the behaviour of a representative customer. Draw $\mu^i \sim log\text{-}MVN_K(\theta_{\mu}, \Sigma_{\mu})$.
```{r}
# calculate gamma and inv wishart params from means and SDs
gammaShRaFromMeanSD <- function( mean , sd ) {
  if ( mean <=0 ) stop("mean must be > 0")
  if ( sd <=0 ) stop("sd must be > 0")
  shape = mean^2/sd^2
  rate = mean/sd^2
  return( list( shape=shape , rate=rate ) )
}

invParamsWishartFromMeanPAndDF <- function(Sigma_mu_mean, df, dim){
  scale_matrix <- Sigma_mu_mean*df - Sigma_mu_mean*dim - Sigma_mu_mean
  return(list(scale_matrix=scale_matrix, df=df, dim=dim))
}

draw_params <- function() {
  alpha <- matrix(nrow = nrow(alpha_mean), ncol = ncol(alpha_mean))
  for(j in c(1:nrow(alpha_mean))) {
    for(k in c(1:ncol(alpha_mean))) {
      gam <- gammaShRaFromMeanSD(mean = alpha_mean[j,k], sd = alpha_sd[j,k])
      alpha[j,k] <- rgamma(n=1, shape = gam$shape, rate = gam$rate)
    }
  }
  
  beta <- numeric()
  for(j in c(1:length(beta_mean))) {
      gam <- gammaShRaFromMeanSD(mean = beta_mean[j], sd = beta_sd[j])
      beta[j] <- rgamma(n=1, shape = gam$shape, rate = gam$rate)
  }
  
  # assumption that all psi_ks are independet we can sample with the diagonal of sds, 
  # in appendix they calculate psi with the Identity matrix which assumes independence!
  psi <- mvrnorm(mu = psi_mean, 
                 Sigma = psi_sd^2 * diag(nrow = length(psi_sd)))
  theta_mu <- mvrnorm(mu = theta_mu_mean, 
                     Sigma = theta_mu_sd^2 * diag(nrow = length(theta_mu_sd)))
  
  inv_wishart_params <- invParamsWishartFromMeanPAndDF(Sigma_mu_mean = Sigma_mu_mean, 
                                                 df = inv_df, dim = length(ad_types))
  Sigma_mu <- riwish(v = inv_wishart_params$df, S = inv_wishart_params$scale_matrix)
  mu <- exp(mvrnorm(mu = theta_mu, Sigma = Sigma_mu))
  list(alpha = alpha, beta = beta, psi = psi, 
       theta_mu = theta_mu, Sigma_mu = Sigma_mu, mu = mu)
}
```
2. Simulate point process in $[0,T]$ given $\alpha, \beta, \psi, \mu^i$ and realized type $j_0$ at $t_0=0 \text{,} (j_0 = 1, \dots, K-1)$.
    a. initialize $t = 0, n = 0, n_K = 0, m = \sum_{k=1}^K (\mu^i_k+\alpha_{j_ok})$
    b. Repeat until $t>T$
        i) Simulate $s \sim Exp(m)$
        ii) Set $t = t + s$
        iii) If $t < T$, calculate 
        $$\lambda_k = \mu_k^i \exp(\psi_k n_K) + \alpha_{j_0 k} \exp(- \beta_{j_0} t) + \sum^n_{l=1, j_l \neq K} \alpha_{j_l k} \exp(- \beta_{j_l}(t-t_l))$$
        and let $\lambda = \sum^K_{k=1} \lambda_k$ and generate $U \sim Unif(0,1)$.
          
            1. If $U \leq \frac{\lambda}{m}$, $n=n+1$, $t_n = t$. Simulate $j_n \sim multinomial(1, \lambda_1/\lambda, \dots,\lambda_K/\lambda)$.    
                - If $j_n=K$ then 
                $$ n_K = n_K+1 $$  
                $$ m = \lambda - \mu_k^i \exp(\psi_k (n_K-1)) + \mu_k^i \exp(\psi_k n_K)$$
                - else $$m = \lambda + \alpha_{j_n k}$$
            2. If $U>\frac{\lambda}{m}$ then $$m=\lambda$$
    c. Simulation output is $\{t_1, \dots, t_n\}$ and $\{j_1, \dots, j_n\}$ 
3. Repeat Step 1 and 2 $R$ times.

        
```{r}
T <- 120
```
      
```{r}
simulate_user_clickstream <- function(alpha, beta, psi, theta_mu, Sigma_mu, mu) {
  t = 0; n = 0; n_K = 0; m = c(alpha[j_0,] + mu) %>% sum; j = integer(); t_j = numeric()
  
  hist_intensity_decay <- function(a, b, events, current_time, event_times) {
    result <- numeric(K)
    names(result) <- ad_types
    if(length(events) == 0) return(result)
    
    for(k in 1:K){
      for(idx in 1:length(events)){
        j_l <- events[idx]
        if(j_l != K) { # do not consider purchases (K) here
          result[k] <- result[k] + 
            alpha[j_l, k] * exp(-1 * beta[j_l] * (t - event_times[idx]))
        }
      }
    }
    return(result)
  }

  repeat {
    s <- rexp(n = 1, rate = m)
    t <- t + s
    if(!is.nan(t) && t < T){
      lambda <- mu * exp(psi*n_K) + alpha[j_0, ] * exp(-1 * beta[j_0] * t)
      lambda <- lambda + hist_intensity_decay(alpha, beta, j, t, t_j)
      U <- runif(n = 1)
      if(is.na(m) || is.na(U)  || NA %in% lambda || is.infinite(m) || Inf %in% lambda) {
        print(paste("Value is NA ", c(m, U, lambda)))
      } 
      if(U <= sum(lambda) / m){
        n <- n + 1
        t_j[n] <- t
        j[n] <- which(rmultinom(n = 1, size = 1, 
                                prob = lambda/sum(lambda))[,1] %in% c(1)) 
        if(j[n] == K) { # if purchase
          n_K <- n_K + 1
          m <- sum(lambda - mu * exp(psi * (n_K -1)) + mu * exp(psi * n_K))
        } else {
          m <- sum(lambda + alpha[j[n], ])
        }
      } else {
        m <- sum(lambda)
      }
    }
    # in some cases n_K becomes greater then a few thousand. 
    # hence, m tends so go to Inf as s becomes 0.
    # this breaks the code and so we break at a certain purchase count.
    if(n > 2 && n > 2 * t_j[n]) return(list())
    #if(t > T || n_K >= 10) return(list(timestamp = t_j, event = ad_types[j]))
    if(t > T) return(list(timestamp = t_j, event = ad_types[j]))
  }
}

```
# Simulate click stream data
## Simulate for one user
```{r}
j_0 <- sample(length(ad_types) - 1, size = 1, prob = first_click_prob)
params <- draw_params()
data <- do.call(simulate_user_clickstream, params)
```
$j_0$ is set to type "`r ad_types[j_0]`" with $T=$ `r T`.
```{r , eval=FALSE, results = 'asis', echo=F}
xtable(data.frame(time = data$timestamp, event = data$event), caption = "Simulated click stream")
```
## Simulate for multiple users and plot their click streams.
```{r, eval=FALSE}
sim_cnt <- 5
```
Simulate for `r sim_cnt` users.
```{r}
sim_data <- function(count) {
  users <- data.frame()
  cnt <- 0
  while(cnt < count) {
    j_0 <- sample(length(ad_types) - 1, size = 1, prob = first_click_prob)
    params <- draw_params()
    data <- do.call(simulate_user_clickstream, params)
    if(!is_empty(data$event)) {
      cnt <- cnt + 1
      users <- users %>% 
        rbind(data_frame(user = rep.int(cnt, length(data$event)), 
                         event = data$event, 
                         timestamp=data$timestamp))
    }
  }
  users
}
```

```{r, eval=FALSE}
users <- sim_data(sim_cnt)
ggplot(users, aes(x = timestamp, y = user, 
                  label = event, color = factor(event), 
                  shape = factor(event),
                  xmax = T)) +
  geom_point(size = 4, alpha = .7) +
  geom_hline(aes(yintercept = user), lty="dotted") + 
  xlim(0, T) + 
  ylim(1, sim_cnt)
```
## Simulate 12,000 users
```{r}
data_file <- "sim_data"
if(file.exists(data_file)) {
  load(data_file) 
} else {
  set.seed(1899)
  data <- sim_data(n)
  save(data, file = data_file)
}

```

```{r, eval=FALSE}
data %>% 
  filter(event != 'purchase') %>%
  group_by(event) %>% 
  summarise(n = n()) %>%
  mutate(ratio = n / sum(n)) %>% 
  knitr::kable(caption = "Ad click ratio")
```

```{r, eval=FALSE}
data %>%
  mutate(event_short = recode(event, 'purchase' = 'P', 'other' = 'O', 
                              'search' = 'S', 'display' = 'D')) %>%
  group_by(user) %>%
  # concatenate all event abbr. into one string
  summarise(stream = paste(event_short, collapse = "")) %>% 
  # only include streams until first P
  mutate(stream = stream %>% gsub(x = ., "(P).*","\\1")) %>% 
  # only display click streams with more than 2 events ending with a purchase
  filter(stream %>% nchar() > 2, stream %>% endsWith('P')) %>% 
  group_by(stream) %>%
  # count distinct click patterns
  summarise(n = n()) %>%
  arrange(desc(n)) %>%
  head(n=20) %>%
  knitr::kable(caption = "Most frequent click combinations ending with a purchase")
```

```{r, eval=FALSE}
data %>%
  mutate(event_short = recode(event, 'purchase' = 'P', 'other' = 'O', 
                              'search' = 'S', 'display' = 'D')) %>%
  group_by(user) %>%
  # concatenate all event abbr. into one string
  summarise(stream = paste(event_short, collapse = "")) %>% 
  mutate(ev_cnt = stream %>% nchar()) %>%
  # cut streams longer than 25 chars
  mutate(stream = ifelse(stream %>% nchar() > 25, 
                         stream %>% strtrim(25) %>% paste("...", sep = ""), stream)) %>%
  group_by(stream) %>%
  #summarise(n = nchar()) %>%
  arrange(desc(ev_cnt)) %>%
  head(n=20) %>%
  knitr::kable(caption = "Longest click streams")
```
Most common click streams
```{r, eval=FALSE}
data %>%
  mutate(event_short = recode(event, 'purchase' = 'P', 'other' = 'O', 
                              'search' = 'S', 'display' = 'D')) %>%
  group_by(user) %>%
  # concatenate all event abbr. into one string
  summarise(stream = paste(event_short, collapse = "")) %>% 
  mutate(ev_cnt = stream %>% nchar()) %>%
  # cut streams longer than 25 chars
  group_by(stream) %>%
  summarise(n = n()) %>%
  arrange(desc(n)) %>% 
  head(n=25)
```
Most common click streams with purchases
```{r, eval=FALSE}
data %>%
  mutate(event_short = recode(event, 'purchase' = 'P', 'other' = 'O', 
                              'search' = 'S', 'display' = 'D')) %>%
  group_by(user) %>%
  # concatenate all event abbr. into one string
  summarise(stream = paste(event_short, collapse = "")) %>% 
  filter(grepl('P', stream)) %>% 
  mutate(ev_cnt = stream %>% nchar()) %>%
  # cut streams longer than 25 chars
  group_by(stream) %>%
  summarise(n = n()) %>%
  arrange(desc(n)) %>% 
  head(n=25)
```


```{r}
all_comb <- expand.grid(user = data$user %>% unique, event = ad_types)
event_counts <- data %>%
  group_by(user, event) %>%
  summarise(n = n()) %>% 
  # join all combinations to get zero count for non-present events
  left_join(all_comb, ., by = c('user', 'event')) %>%
  mutate_if(is.integer, funs(replace(., is.na(.), 0)))
rm(all_comb)

```

```{r, eval=FALSE}
event_counts %>%
  ggplot(aes(y = n, x = event)) +
  geom_boxplot()
```
```{r, eval=FALSE}
event_counts %>%
  ggplot(aes(y = n, x = event)) +
  geom_boxplot() + 
  ylim(c(0, 5))
```
Transistion Matrix for 100 users
```{r}
streams <- data %>%
  filter(user %in% 1:user_cnt) %>% 
  mutate(event_short = recode(event, 'purchase' = 'P', 'other' = 'O', 
                              'search' = 'S', 'display' = 'D')) %>%
  group_by(user) %>%
  summarise(stream = paste(event_short, collapse = "")) %>%
  select(stream) %>%
  filter(nchar(stream) > 1) %>%
  .$stream %>% 
  as.vector()

adj_list <- foreach(s = iter(streams), .combine = append) %dopar% {
  substring(s, 1:(nchar(s) - 1), 2:nchar(s)) %>% str_split(pattern = "")
}

adj_mat <- matrix(rep(0, K * K), 
                  ncol = K, 
                  dimnames = list(c("S", "D", "O", "P"), c("S", "D", "O", "P"))) 
 
for(i in adj_list) {
  adj_mat[i[1], i[2]] <- adj_mat[i[1], i[2]] + 1
}
rownames(adj_mat) <- ad_types
colnames(adj_mat) <- ad_types
adj_mat %>% xtable(caption = "Transition matrix for events related to 100 users.", label = "tab_trans_100", digits = 0)
```
Transistion Matrix for 500 users
```{r}
streams <- data %>%
  mutate(event_short = recode(event, 'purchase' = 'P', 'other' = 'O', 
                              'search' = 'S', 'display' = 'D')) %>%
  group_by(user) %>%
  summarise(stream = paste(event_short, collapse = "")) %>%
  head(500) %>% 
  select(stream) %>%
  filter(nchar(stream) > 1) %>%
  .$stream %>% 
  as.vector()

adj_list <- foreach(s = iter(streams), .combine = append) %dopar% {
  substring(s, 1:(nchar(s) - 1), 2:nchar(s)) %>% str_split(pattern = "")
}

adj_mat <- matrix(rep(0, K * K), 
                  ncol = K, 
                  dimnames = list(c("S", "D", "O", "P"), c("S", "D", "O", "P"))) 
 
for(i in adj_list) {
  adj_mat[i[1], i[2]] <- adj_mat[i[1], i[2]] + 1
}
rownames(adj_mat) <- ad_types
colnames(adj_mat) <- ad_types
adj_mat %>% xtable(caption = "Transition matrix for events related to 500 users.", label = "tab_trans_500", digits = 0)
```

## Bring data into correct format
```{r}
data_new <- data %>% group_by(user) %>% mutate(search=ave(event == "search", FUN=cumsum))

data_new <- data_new %>% group_by(user) %>% mutate(display=ave(event == "display", FUN=cumsum))

data_new <- data_new %>% group_by(user) %>% mutate(other=ave(event == "other", FUN=cumsum))

data_new <- data_new %>% group_by(user) %>% mutate(purchase=ave(event == "purchase", FUN=cumsum))

data_new <- data_new %>% ungroup %>% as.data.frame

total_counts <- data %>% group_by(user, event) %>% mutate(total_counts = n()) %>% select(user, event, total_counts) %>% distinct %>% ungroup

total_counts <- total_counts %>% spread(event, total_counts) # bring to wide format
total_counts[is.na(total_counts)] <- 0
# reorder columns
total_counts <- total_counts[, c(1,5,2,3,4)]
total_counts <- as.data.frame(total_counts)
data_new$event <- as.factor(data_new$event)
levels(data_new$event) <- c(2,3,4,1)
```


# Log Likelihood

```{r}
likelihood <- function(i, params){
  llAlpha <- params$alpha
  llBeta <- params$beta
  llPsi <- params$psi
  llMu <- params$mu[i,]
  data_new_i <- data_new %>% filter(user == i)
  sum_k_log <- 0 # wird spaeter mit ergebnis aus k-loop multipliziert
  for (k in 1:length(ad_types)){
    prod_l <- 1 # wird sp?ter mit ergebnis aus l-loop multipliziert
    if (total_counts[i , k+1] != 0){
      first_two_lines <- numeric()
      for (l in 1:(total_counts[i , k+1])){ 
        first_line <- llMu[k] * exp(llPsi[k] * (data_new_i %>% 
                                              filter(event == k) %>%
                                                  .[l, ] %>% 
                                                  select(purchase))) %>% 
          as.numeric() # erste Zeile
        
        sum_j <- numeric()
        for (j in 1:3){ # zweite Zeile erste summe 1:K-1
          # j schleife umgehen
            sum_index <- ifelse(data_new_i %>% 
                                  filter(event == k) %>% 
                                  plyr::empty(), #if is empty
                                0, #then
                                data_new_i %>% #else
                                  filter(event == k) %>% 
                                  filter(row_number() ==l) %>% 
                                  select(j+3) %>% as.numeric()
                                )
            if (sum_index > 0){
            sum_m <- numeric()
            t_l <- data_new_i %>% 
              filter(event==k) %>% 
              filter(row_number() == l) %>% 
              select(timestamp) %>% as.numeric()
            t_m <- data_new_i %>% filter(event==j) %>% 
              select(timestamp) 
            # second line:
            sum_m <- llAlpha[j,k] * exp(-llBeta[j] * (t_l - t_m[1:sum_index,]))
              
              } else {sum_m <- 0} 
            sum_j[j] <- sum(sum_m) # Summe j
            } # j Ende
        first_two_lines[l] <- first_line + sum(sum_j) # erste und zweite Zeile
      } 
    } else {
      first_two_lines <- 1
    }
    prod_l <- prod(first_two_lines)
    
      # Produkt ueber l; wenn von 1 bis 0, dann ergebnis auf 0 setzen
    sum_m <- 0
    purchase_cnt <- total_counts[i,5]
    if (purchase_cnt > 0){
    for (m in 0:purchase_cnt){
      # 0 bis Anzahl Kaeufe in 120 Tagen -1
      if(data_new_i %>% filter(event == K) %>% 
                         nrow() >= m+1){
        
        t_m1 <- data_new_i %>% filter(event == K) %>% 
                           filter(row_number() == m+1) %>% 
                           select(timestamp) %>% 
                           as.numeric()
        t_m <- ifelse(m == 0, 
                      0,
                       data_new_i %>% filter(event == K) %>% 
                            filter(row_number() == m) %>% 
                            select(timestamp) %>% as.numeric()
      )
      tmp_m <- exp(llPsi[k]*m)*(t_m1 - t_m)
      } else {tmp_m <- 0}
      sum_m <- sum_m + tmp_m # Summe m
    } # m Ende
    exponent_teil1 <- sum_m * -llMu[k]
    
    } # if ende
    else {exponent_teil1 <- 0}
    browser(expr = is.na(exponent_teil1))
    exponent_teil2 <- 0
    for (j in 1:3){ # letzte Zeile erste Summe
      # j schleife umgehen?
      sum_m <- numeric()
      if (total_counts[i,j+1] > 0) {
        #for (m in 1:total_counts[i,j+1]){ # letzte Zeile zweite Summe
          t_m_neu <- data_new_i %>% filter(event == j) %>% 
            #filter(row_number() == m) %>% 
            select(timestamp) # %>% as.numeric()
          tmp_m <- (llAlpha[j,k]/llBeta[j]) * (1-exp(-llBeta[j] * (T - t_m_neu)))
          sum_m <- sum(tmp_m) # Summe m
        #} # m Ende
        exponent_teil2 <- exponent_teil2 + sum_m
      } 
    } # j Ende
    
    third_fourth_line <- exponent_teil1 - exponent_teil2
    sum_k_log <- log(prod_l) + third_fourth_line + sum_k_log %>% as.numeric()
  }  # loop over all ad types
  return(sum_k_log)
} # Funktion Ende
```



## 500 Users
Run MCMC or load eisting results
```{r}
if(file.exists('mcmc_trajectory_bak_500u_1000i')) {
  load('results/mcmc_trajectory_bak_500u_1000i')
} else {
  source('mcmc.R')
}
```
Estimated $\alpha$
```{r}
burnIn <- 0.4 * length(trajectory)
alpha_estimated <- foreach(a = iter(trajectory[burnIn:length(trajectory)]), 
                           .combine = '+') %dopar% {a$alpha} / length(trajectory)

rownames(alpha_estimated) <- ad_types[1:K-1]
colnames(alpha_estimated) <- ad_types
alpha_estimated %>% xtable(caption= "Estimated $\\alpha$ parameters for 500 users after 1000 iterations with burn rate of 0.4", label = "tab_alpha_est_500", display = c('s', rep('e', K)))
```
Estimated $\theta_{\mu}$
```{r}
theta_mu_estimated <- foreach(t = iter(trajectory[burnIn:length(trajectory)]), .combine = rbind) %dopar% 
  t$theta_mu %>% colMeans %>% matrix(ncol = K)

colnames(theta_mu_estimated) <- ad_types
theta_mu_estimated  %>% 
  xtable(caption="Estimated $\\theta_{\\mu}$ after 1000 iteration with burn in rate of 0.4",
                              label = "tab_theta_est_500", digits = 5, display = c('s',rep('g', K)))
```
Estimated $\psi$
```{r}
psi_estimated <- foreach(t = iter(trajectory[burnIn:length(trajectory)]), .combine = rbind) %dopar% 
  t$psi %>% colMeans %>% matrix(ncol = K)

colnames(psi_estimated) <- ad_types
psi_estimated %>% xtable(caption = "Estimated $\\psi$ for 500 users after 1000 iteration with burn in rate of 0.4.", digits = 5, display = c('s',rep('g', K)), label = "tab_psi_est_500")
```
Estimated $\beta$
```{r}
beta_estimated <- foreach(t = iter(trajectory[burnIn:length(trajectory)]), .combine = rbind) %dopar% 
  t$beta %>% colMeans %>% matrix(ncol = K-1)

colnames(beta_estimated) <- ad_types[1:K-1]

beta_estimated %>% xtable(caption = "Estimated $\\beta$ for 500 users after 1000 iteration with burn in rate of 0.4.", digits = 5, display = c('s',rep('g', K-1)), label = "tab_beta_est_500")
```
Estimated $\Sigma_{\mu}$

```{r}
Sigma_estimated <- foreach(a = iter(trajectory[burnIn:length(trajectory)]), 
                           .combine = '+') %dopar% {a$Sigma_mu} / length(trajectory)

rownames(Sigma_estimated) <- ad_types
colnames(Sigma_estimated) <- ad_types
Sigma_estimated %>% xtable(caption= "Estimated $\\Sigma_{\\mu}$ parameters for 500 users after 1000 iterations with burn rate of 0.4", label = "tab_sigma_est_500", display = c('s', rep('e', K)))
```
### MCMC Diagnostics
```{r}
beta_mcmc <- foreach(t = iter(trajectory[burnIn:length(trajectory)]), .combine = rbind) %dopar% 
  t$beta %>% matrix(ncol = K-1) %>% as.mcmc
traceplot(beta_mcmc)
```
```{r}
densplot(beta_mcmc)
```
```{r}
HPDinterval(beta_mcmc)
```
```{r}
theta_mcmc <- foreach(t = iter(trajectory[burnIn:length(trajectory)]), .combine = rbind) %dopar% 
  t$theta_mu %>% matrix(ncol = K) %>% as.mcmc
traceplot(theta_mcmc)
```
```{r}
HDIofMCMC(theta_mcmc)
```
```{r}
densplot(theta_mcmc)
```

```{r}
psi_mcmc <- foreach(t = iter(trajectory[burnIn:length(trajectory)]), .combine = rbind) %dopar% 
  t$psi %>% matrix(ncol = K) %>% as.mcmc
traceplot(theta_mcmc)
```
```{r}
densplot(psi_mcmc)
```


```{r}
acceptedTraj <- trajectory[ (burnIn+1) : length(trajectory) ]

alphaMcmc <- foreach(t = iter(trajectory[burnIn:length(trajectory)]), .combine = rbind) %dopar% 
  {t$alpha %>% as.vector()} %>% matrix(ncol = K * (K-1)) %>% as.mcmc
traceplot(alphaMcmc)
```

```{r}
densplot(alphaMcmc)
```
## 100 Users
Run MCMC or load eisting results
```{r}
if(file.exists('results/mcmc_trajectory_bak_100u_2000i')) {
  load('results/mcmc_trajectory_bak_100u_2000i')
} else {
  source('mcmc.R')
}
```
Estimated $\alpha$
```{r}
burnIn <- 0.25 * length(trajectory)
alpha_estimated <- foreach(a = iter(trajectory[burnIn:length(trajectory)]), 
                           .combine = '+') %dopar% {a$alpha} / length(trajectory)

rownames(alpha_estimated) <- ad_types[1:K-1]
colnames(alpha_estimated) <- ad_types
alpha_estimated %>% xtable(caption= "Estimated $\\alpha$ parameters for 100 users after 2000 iterations with burn rate of 0.25", label = "tab_alpha_est_100", display = c('s', rep('e', K)))
```
Estimated $\theta_{\mu}$
```{r}
theta_mu_estimated <- foreach(t = iter(trajectory[burnIn:length(trajectory)]), .combine = rbind) %dopar% 
  t$theta_mu %>% colMeans %>% matrix(ncol = K)

colnames(theta_mu_estimated) <- ad_types
theta_mu_estimated  %>% 
  xtable(caption="Estimated $\\theta_{\\mu}$ after 2000 iteration with burn in rate of 0.25.",
                              label = "tab_theta_est_100", digits = 5, display = c('s',rep('g', K)))
```
Estimated $\psi$
```{r}
psi_estimated <- foreach(t = iter(trajectory[burnIn:length(trajectory)]), .combine = rbind) %dopar% 
  t$psi %>% colMeans %>% matrix(ncol = K)

colnames(psi_estimated) <- ad_types
psi_estimated %>% xtable(caption = "Estimated $\\psi$ for 100 users after 2000 iteration with burn in rate of 0.25.", digits = 5, display = c('s',rep('g', K)), label = "tab_psi_est_100")
```
Estimated $\beta$
```{r}
beta_estimated <- foreach(t = iter(trajectory[burnIn:length(trajectory)]), .combine = rbind) %dopar% 
  t$beta %>% colMeans %>% matrix(ncol = K-1)

colnames(beta_estimated) <- ad_types[1:K-1]

beta_estimated %>% xtable(caption = "Estimated $\\beta$ for 100 users after 2000 iteration with burn in rate of 0.25.", digits = 5, display = c('s',rep('g', K-1)), label = "tab_beta_est_100")
```
Estimated $\Sigma_{\mu}$

```{r}
Sigma_estimated <- foreach(a = iter(trajectory[burnIn:length(trajectory)]), 
                           .combine = '+') %dopar% {a$Sigma_mu} / length(trajectory)

rownames(Sigma_estimated) <- ad_types
colnames(Sigma_estimated) <- ad_types
Sigma_estimated %>% xtable(caption= "Estimated $\\Sigma_{\\mu}$ parameters for 100 users after 2000 iterations with burn rate of 0.25", label = "tab_sigma_est_100", display = c('s', rep('e', K)))
```
## MCMC Diagnostics
```{r}
beta_mcmc <- foreach(t = iter(trajectory[burnIn:length(trajectory)]), .combine = rbind) %dopar% 
  t$beta %>% matrix(ncol = K-1) %>% as.mcmc
traceplot(beta_mcmc)
```
```{r}
densplot(beta_mcmc)
```
```{r}
HPDinterval(beta_mcmc)
```
```{r}
theta_mcmc <- foreach(t = iter(trajectory[burnIn:length(trajectory)]), .combine = rbind) %dopar% 
  t$theta_mu %>% matrix(ncol = K) %>% as.mcmc
traceplot(theta_mcmc)
```
```{r}
HDIofMCMC(theta_mcmc)
```
```{r}
densplot(theta_mcmc)
```

```{r}
psi_mcmc <- foreach(t = iter(trajectory[burnIn:length(trajectory)]), .combine = rbind) %dopar% 
  t$psi %>% matrix(ncol = K) %>% as.mcmc
traceplot(theta_mcmc)
```
```{r}
densplot(psi_mcmc)
```


```{r}
acceptedTraj <- trajectory[ (burnIn+1) : length(trajectory) ]

alphaMcmc <- foreach(t = iter(trajectory[burnIn:length(trajectory)]), .combine = rbind) %dopar% 
  {t$alpha %>% as.vector()} %>% matrix(ncol = K * (K-1)) %>% as.mcmc
traceplot(alphaMcmc)
```

```{r}
densplot(alphaMcmc)
```
